Supplementary Information

### 2021-06-10 doc/manuscript/ForestGEO_treerings.Rmd
google docs: misc cleanup
citations problem #133

#SI table captions
site_details_legend <- "Table S1. Site Details."
species_table_legend <-"Table S2. Species analyzed, their characteristics, and bark allometries applied."
sampling_details_legend <- "Table S3. Sampling details for species by site."
bark_allometries_legend <- "Table S4. Allometric equations for bark thickness."
qualitative_results_comparison_legend <- "Table S5. Qualtiative comparison of results from this study with previous studies employing conventional methods."

@@

In all seven sites dominated by trees less than 100 years old, growth trends were universally negative, when significant (Table `r hypotheses`, Fig. `r year_fig`). 
While attribution of these trends is difficult, it is likely that some trends are caused by stand dynamics as cohorts and stands develop over time [@oliver_forest_1990]. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., *Acer rubrum* and *Betula alleghaniensis* at HF), persistent differences in growth rates among individuals could produce a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017; see also @vandersleen_trends_2017].
Particularly in secondary stands where many of the sampled species recruited in pulses that were followed by low recruitment (e.g., SCBI, HF; Appendix `r site_details_appendix`), growth declines are consistent with the tendency for faster tree growth during early stand development [@lorimer_estimating_1988; @lorimer_methodology_1989; @oliver_forest_1990], and with increasing competition and declining woody productivity as young stands mature [e.g., @goulden_patterns_2011; @pregitzer_carbon_2004; @west_increasing_2020].
Abiotic drivers likely also play a role at some of these sites.
At Scotty Creek, in northern Canada, rapid warming is thawing permafrost and altering hydrologic conditions [@baltzer_forests_2014], resulting in high mortality, growth declines, and low recruitment of *Picea mariana* [@sniderhan_growth_2016; @dearborn_permafrost_2020]; we attribute pronounced negative growth trends to a combination of successional declines and indirect climatic stress.

@@

# Authors' contributions {-#Authors}
KAT, VH, CR, RA, CP conceived the ideas and designed methodology; 
NP, CDA, RAS, TA, JLB, JDB, SB, PC, RH, JK, JL, EQM, JM, PS, AES, AJT, IV, MV, and PAZ collected the data; 
VH, BG, EGA, CD, and NP organized and analysed the data; 
KAT led the writing of the manuscript. 
All authors contributed critically to the drafts and gave final approval for publication.

>>>>

#SI table captions
site_details_legend <- "Table S1. Site Details."
species_table_legend <-"Table S2. Species analyzed, their characteristics, and bark allometries applied."
sampling_details_legend <- "Table S3. Sampling details for species by site."
bark_allometries_legend <- "Table S4. Allometric equations for bark thickness."
qualitative_results_comparison_legend <- "Table S5. Qualitative comparison of results from this study with previous studies employing conventional methods."

@@

In all seven sites dominated by trees less than 100 years old, growth trends were universally negative, when significant (Table `r hypotheses`, Fig. `r year_fig`). 
While attribution of these trends is difficult, it is likely that some trends are caused by stand dynamics as cohorts and stands develop over time.
Such negative trends are fairly typical of mixed-species stands that experience vertical stratification [@oliver_forest_1990]. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., *Acer rubrum* and *Betula alleghaniensis* at HF), persistent differences in growth rates among individuals could produce a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017; see also @vandersleen_trends_2017].
Particularly in secondary stands where many of the sampled species recruited in pulses that were followed by low recruitment (e.g., SCBI, HF; Appendix `r site_details_appendix`), growth declines are consistent with the tendency for faster tree growth during early stand development [@lorimer_estimating_1988; @lorimer_methodology_1989; @oliver_forest_1990], and with increasing competition and declining woody productivity as young stands mature [e.g., @goulden_patterns_2011; @pregitzer_carbon_2004; @west_increasing_2020].
Gradual shifts in abiotic drivers (e.g., indirect effects of warming) likely also play a role at some of these sites.
At Scotty Creek, in northern Canada, rapid warming is thawing permafrost and altering hydrologic conditions [@baltzer_forests_2014], resulting in high mortality, growth declines, and low recruitment of *Picea mariana* [@sniderhan_growth_2016; @dearborn_permafrost_2020]. At this site, we attribute pronounced negative growth trends to a combination of successional declines and indirect climatic stress.

@@

# Authors' contributions {-#Authors}
KAT, VH, CR, MRA, CP conceived the ideas and designed methodology; 
NP, CDA, RAS, TA, JLB, JDB, SB, PC, RH, JK, JL, EQM, JTM, PS, AES, AJT, IV, MV, and PAZ collected the data; 
VH, BG, EGA, CD, and NP organized and analysed the data; 
KAT led the writing of the manuscript. 
All authors contributed critically to the drafts and gave final approval for publication.
